---
title: "All about you!"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
    css: "css/app-style.css"
runtime: shiny_prerendered
---


<!-- Add JavaScript code for making the exercise code larger -->
<script language="JavaScript" src="js/exercise-font-size.js"></script>


```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(tidyverse)
library(googlesheets4)
library(viridis)
library(waffle)
library(urbnmapr)


# remove authorization from googlesheet
gs4_deauth()

tutorial_options(
  # code running in exercise times out after 30 seconds
  exercise.timelimit = 30, 
  )

# hide non-exercise code chunks
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE, 
                      message = FALSE)

# set ggplot options 

theme_set(theme_bw())

viridis_qualitative_pal7 <- c("#440154FF", "#DCE319FF", "#443A83FF", 
                              "#8FD744FF", "#31688EFF",  "#35B779FF", 
                              "#21908CFF") #replace the first yellow color

scale_fill_discrete <- function(...) {
   scale_fill_manual(..., values = viridis_qualitative_pal7)
} 

scale_colour_discrete <- function(...) {
  scale_colour_manual(..., values = viridis_qualitative_pal7)
}


# read in survey responses
responses <- read_sheet("https://docs.google.com/spreadsheets/d/1zjoOj4UoXwzuuuJRQCkTd8eeRIjAMWXSmomR8e2QT8c/edit?usp=sharing")

# prep data 
responses <- responses %>%
  select(-Timestamp) %>%
  rename(state = `What state do you currently live in?`,
         school_level = `Are you a high school or undergraduate student?`,
         num_classes = `How many classes are you currently taking?`,
         device_school = `About what percent of waking hours do you spend using an electronic device for something related to school, such as attending classes, studying, homework, etc.? Enter the value as a number 0 - 100%.`, 
         device_leisure = `About what percent of waking hours do you spend on an electronic device for leisure activities, such as social media, watching TV, chatting with friends, etc.? Enter the value as a number 0 - 100%.`, 
         social_media = `What social media platforms do you use? Select all that apply.`)
```

## Introduction 

Welcome! In this tutorial, we'll learn R functions that can help us wrangle, summarize, and visualize your survey responses and learn a little more about you! We'll build upon what you learned in the first tutorial and explore the relationships between multiple variables.

### Packages

We'll use the **tidyverse**, a collection of R packages designed for data science.

::: {.your-turn}
Click *Run Code* to load the tidyverse package.
:::

```{r load-packages, exercise = TRUE}
library(tidyverse)
```

## Data

### Your survey responses 

The data frame `responses` contains your responses to the [survey](https://forms.gle/2VhEZPqEjeN4ykY87) you took during this workshop. 

Here is the codebook for `responses`:

- `state`: state you live in
- `school_level`: high school or undergraduate student
- `num_classes`: number of classes you're currently taking
- `device_school`: percent of waking hours spent using electronic devices for school 
- `device_leisure`: percent of waking hours spent using electronic devices for leisure activities
- `social_media`: social media platforms you typically use (Facebook, Twitter, Instagram, Snapchat, TikTok, Other, None/ I don't use social media)

### Let's take look at the data

::: {.your-turn}
Click run **Run Code** to view the data!
::: 

```{r view-data, exercise = TRUE}
responses
```

## School

### School level

**How many students are high school students? College undergraduates?** To answer this question, we can use the `count` function to count the number of students at each level of `school_level`. 

```{r school-level-count, exercise = TRUE}
responses %>%
  count(___)
```

Now let's make a bar plot to visualize these counts! 

```{r school-level-bar, exercise  = TRUE}
ggplot(data = responses, aes(x = school_level)) +
  geom_bar() 
```

### Customize the bar plot

There are many ways to customize plots made with `ggplot`. For now, let's change the color of the bars and add a title. 

- To change the color of the bars, we'll include the argument `fill = "[COLOR]"` inside `geom_bar()`. In the code below, you can see that we've changed the color of the bars to blue. 
- We'll add a title using `labs()`. 

::: {.your-turn}
Update the code by filling in a color of your choice and adding a title. You can [click here](https://www.r-graph-gallery.com/42-colors-names.html) for a list of colors. 
::: 

```{r school-level-custom, exercise = TRUE}
ggplot(data = responses, aes(x = school_level)) +
  geom_bar(fill = "___") + 
  labs(title = "___")
```


## Electronic devices

**What's the relationship between the percent of time students use an electronic device for school (`device_school`) versus for leisure activities (`device_leisure`)?**

**Let's make a scatterplot to find out!**

### Electronic device use

We can use `geom_point()` to make scatterplots with `ggplot`. 

```{r time-on-device, exercise = TRUE}
ggplot(data = responses, aes(x = device_school, y = device_leisure)) +
  geom_point()
```


### Visualize general trend

We can also use `geom_smooth()` to get a better view of the general relationship between the variables.  

```{r device-by-format-smooth, exercise = TRUE}
ggplot(data = responses, aes(x = device_school, y = device_leisure)) + 
  geom_point() +
  geom_smooth()
```

### Add labels and a title!

```{r device-by-format-facet-2, exercise = TRUE}
ggplot(data = responses, aes(x = device_school, y = device_leisure)) +
  geom_point() +
  geom_smooth() +
  labs(x = "___", 
      y = "___", 
      title = "Electronic device use for school and leisure")
```


**What do you observe about the relationship between `device_school` and `device_leisure`?**


### Your turn!

Now let's see if the relationship between `device_school` and `device_leisure` differs for high school and undergraduate students. 

::: {.your-turn}
Make a scatterplot the relationship between percent of time using an electronic device for school (`device_school`) and for leisure activities (`device_leisure`) coloring the  points by `school_level`.
:::

```{r device-by-region, exercise = TRUE}
ggplot(data = responses, aes(x = ___, y = ___, color = ___)) +
  geom_point() +
  geom_smooth() +
  labs(x = "___", 
      y = "___", 
      color = "___"
      title = "___")
```

**What do you observe from your plot?** 

## Mapping states

**Let's take a look at which states are represented today!** 

We'll start by counting the number of students who live in each state. We'll use the `arrange()` function so we can see the states with the most students at the top of the list. 

```{r state-count, exercise = TRUE}
responses %>%
  ___(___) %>%
  arrange(desc(n))
```


### Map of the United States

We can use R to plot the number of students from each state onto a map. We'll start by looking at a map of the United States. The data set `usa_map_data` contains the coordinates needed to create a map of the US, and the function `geom_polygon()`  uses the coordinates to produce a map. 


```{r map-data, include = F}
usa_map_data <- urbnmapr::states %>%
  rename(state = state_name)
```


```{r map, exercise = TRUE, exercise.setup = "map-data"}
ggplot(data = usa_map_data, aes(x = long, y = lat, group = group)) + 
  geom_polygon()
```

### Customize the map

::: {.your-turn}
Fill in the blanks with colors of your choice to customize your map. Notice how the map changes for different values of `fill` versus different values of `color`. 
:::

```{r map-customize, exercise = TRUE, exercise.setup = "map-data"}
ggplot(data = usa_map_data, aes(x = long, y = lat, group = group)) + 
  geom_polygon(fill = "___", color = "___")
```


### Color states by number of students

Instead of using the same color for all states, let's color in the the states based on the number of students who live in that state. 

To do so we will `join` the count of the number of students from each state to `usa_map_data`. Use `<-` to save the results in a new data frame (a *data frame* is the name for a data set in R).

```{r join-data, exercise = TRUE, exercise.setup = "map-data"}
num_students <- responses %>%
  count(state)

usa_map_data <- usa_map_data %>%
  left_join(num_students, by = c("state"))
```

```{r joined-map-data, include = F}
num_students <- responses %>%
  count(state)

usa_map_data <- usa_map_data %>%
  left_join(num_students, by = c("state"))
```

Now fill in the color of the states based on the number of students who live in that state. Add a title to the map. 

```{r map-by-num, exercise = TRUE, exercise.setup = "joined-map-data"}
ggplot(data = usa_map_data, aes(x = long, y = lat, group = group, fill = ___)) + 
  geom_polygon(color = "white") + 
  labs(title = "____")
```

**What do you observe from the plot?**

## Social media

### View `social_media`

Now let's analyze the social media data. Our ultimate goal is to answer the following question: 

**Are students who use Instagram more likely to use TikTok than students who don't use Instagram?**

Let's start by using the `select` function to take a look at the social media variable.

```{r instagram-1, exercise = TRUE}
responses %>%
  select(social_media)
```

We want to calculate the percent of students who use a particular social media platform. 
Students could select multiple social media platforms in the survey, so we need to make a new variable to indicate whether the platform we're analyzing is included in the response.

### Make Instagram variable

Let's start by looking at Instagram. We'll use the `str_detect` and `mutate` functions to make a new variable that indicates whether a student uses Instagram.

- `str_detect`: Returns the value **TRUE** if "Instagram" is included in a response and  **FALSE** otherwise.
- `mutate`: Add the new variable to the data frame


Run the code below to make the new variable `instagram`.

```{r instagram-2, exercise = TRUE}
responses <- responses %>%
  mutate(instagram  = str_detect(social_media, "Instagram"))
```

Now let's take a look at our new variable! 

```{r instagram-view, exercise = TRUE, exercise.setup = "instagram-2"}
responses %>%
  select(social_media, instagram)
```

Make a barplot to visualize the number of students who do and do not use Instagram. 

```{r instagram-bar, exercise  = TRUE}
ggplot(data = responses, aes(x = instagram)) +
  geom_bar() 
```



### Calculate percent Instagram 

We want to calculate the percent of students who use Instagram. We'll start by counting the number of students in each category of `instagram` then create a new variable of the corresponding percents. 

```{r instagram-percent, exercise = TRUE, exercise.setup = "instagram-2"}
responses %>%
  count(___) %>%
  mutate(percent = n / sum(n) * 100)
```


### Your turn! TikTok users

What percent of students use TikTok? Make a new variable that indicates whether a student selected TikTok, then calculate the percent of students who are TikTok users.

::: {.your-turn}
Create a variable to indicate whether students use TikTok, then view the data with your new variable. 
::: 

```{r tik-tok, exercise = TRUE, exercise.setup = "instagram-2"}
responses <- responses %>%
  ___(tiktok = ___(social_media, "___"))

responses %>%
  select(social_media, tiktok)
```

::: {.your-turn}
Calculate the percent of students who use TikTok! 
::: 

```{r tik-tok-data, include  = FALSE}
responses <- responses %>%
  mutate(tiktok = str_detect(social_media, "TikTok"), 
         instagram = str_detect(social_media, "Instagram"))
```

```{r tik-tok-percent, exercise = TRUE, exercise.setup = "tik-tok-data"}
responses %>%
  ___(___) %>%
  mutate(percent = n / sum(___) * 100)
```

### Relationship between using TikTok and Instagram

**Are students who use Instagram more likely to use TikTok?** 

::: {.your-turn}
Calculate the percent of TikTok users for each level of `instagram`.
::: 

```{r tik-tok-instagram, exercise = TRUE, exercise.setup = "tik-tok-data" }
responses %>%
  
```

```{r tik-tok-instagram-hint}
responses %>%
  group_by(___) %>%
  ___(___) %>%
  ___(percent = ___ / sum(___) * 100)
```


Let's visualize the relationship between Instagram and TikTok use. We can update the original bar plot of `instgram` by filling in the color of the bars based on `tiktok`. This is a **stacked barplot**.  

Since we want to fill in the color of the bars based on a variable (instead of making them the same color as before), we will 

- Define `fill = [VAR]` as an argument in `aes()`. 
- Include `position = "fill"` in `geom_bar()`


::: {.your-turn}
Make a stacked bar plot to view the relationship between the two variables! 
:::

```{r tik-tok-instagram-2, exercise = TRUE, exercise.setup = "tik-tok-data" }
ggplot(data = responses, aes(x = ___, fill = ___)) +
  geom_bar(position = "___") 
```


**What is the relationship between using Instagram and TikTok? Are students who use Instagram more likely to use TikTok?**

## Explore!

**Use the space below to explore the data, calculate new statistics, and create new visualizations.**


```{r explore, exercise = TRUE, exercise.setup = "tik-tok-data" }

```



## Learn more!

Below are a few resources to keep learning more about R and the tidyverse!

- [R for Data Science](https://r4ds.had.co.nz/)
- [RStudio Primers](https://rstudio.cloud/learn/primers)
- [tidyverse.org](https://www.tidyverse.org/)



