---
title: "TikTok Lyrics"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
    css: "css/font-size.css"
runtime: shiny_prerendered
---

<!-- Add JavaScript code for making the exercise code larger -->
<script language="JavaScript" src="js/exercise-font-size.js"></script>


```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(tidytext)

lyrics <- read_rds(here::here("tiktok-lyrics", "data/lyrics.rds"))
lyrics_words <- read_rds(here::here("tiktok-lyrics", "data/lyric-words.rds"))
```

## Introduction

### 

```{r guess-1}
question_text(
  "Can you tell what the following visualisation (that is missing labels) might be based on?",
  answer("Frequently used words in songs from 2020-2021 viral TikTok videos", correct = TRUE),
  correct = "You'll find out shortly whether your guess is correct!",
  incorrect = "You'll find out shortly whether your guess is correct!"
  )
```

```{r out.width = "100%", fig.width = 10, fig.asp = 0.7}
lyrics_words %>%
  count(track_title, word, sort = TRUE) %>%
  group_by(track_title) %>%
  slice_head(n = 10) %>%
  ungroup() %>%
  ggplot(aes(y = reorder_within(word, n, track_title), x = n, fill = track_title)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(vars(track_title), ncol = 3, scales = "free") +
  scale_y_reordered() +
  scale_fill_viridis_d(option = "E") +
  theme_minimal(base_size = 14) +
  labs(
    x = NULL,
    y = NULL
  ) +
  theme(
    strip.text = element_blank(),
    axis.text.x = element_blank()
    )
```

###

```{r guess-2}
question_text(
  "How about now? Can you tell what the following visualisation (that is missing labels) might be based on?",
  answer("Frequently used words in songs from 2020-2021 viral TikTok videos", correct = TRUE),
  correct = "You'll find out shortly whether your guess is correct!",
  incorrect = "You'll find out shortly whether your guess is correct!"
  )
```

```{r out.width = "100%", fig.width = 10, fig.asp = 0.7}
lyrics_words %>%
  count(track_title, word, sort = TRUE) %>%
  group_by(track_title) %>%
  slice_head(n = 10) %>%
  ungroup() %>%
  ggplot(aes(y = reorder_within(word, n, track_title), x = n, fill = track_title)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(vars(track_title), ncol = 3, scales = "free") +
  scale_y_reordered() +
  scale_fill_viridis_d(option = "E") +
  theme_minimal(base_size = 14) +
  labs(
    x = "Frequency",
    y = "Word",
    title = "Lyrics"
  ) +
  theme(
    strip.text = element_blank(),
    axis.text.x = element_blank()
    )
```

###

Let's reveal the answer!

```{r out.width = "100%", fig.width = 10, fig.asp = 0.8}
lyrics_words %>%
  count(track_title, word, sort = TRUE) %>%
  group_by(track_title) %>%
  slice_head(n = 10) %>%
  ungroup() %>%
  ggplot(aes(y = reorder_within(word, n, track_title), x = n, fill = track_title)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(vars(track_title), ncol = 3, scales = "free") +
  scale_y_reordered() +
  scale_fill_viridis_d(option = "E") +
  theme_minimal(base_size = 14) +
  labs(
    x = "Frequency",
    y = NULL,
    title = "Frequently used words in songs from 2020-2021 viral TikTok videos",
    subtitle = "After removing stop words and profanity",
    caption = "Data: Genius - Song Lyrics & Knowledge"
  )
```



```{r}
knitr::knit_exit()
```

## Topic 1

### Exercise 

*Here's a simple exercise with an empty code chunk provided for entering the answer.*

Write the R code required to add two plus two:

```{r two-plus-two, exercise=TRUE}

```

### Exercise with Code

*Here's an exercise with some prepopulated code as well as `exercise.lines = 5` to provide a bit more initial room to work.*

Now write a function that adds any two numbers and then call it:

```{r add-function, exercise=TRUE, exercise.lines = 5}
add <- function() {
  
}
```

## Topic 2

### Exercise with Hint

*Here's an exercise where the chunk is pre-evaluated via the `exercise.eval` option (so the user can see the default output we'd like them to customize). We also add a "hint" to the correct solution via the chunk immediate below labeled `print-limit-hint`.*

Modify the following code to limit the number of rows printed to 5:

```{r print-limit, exercise=TRUE, exercise.eval=TRUE}
mtcars
```

```{r print-limit-hint}
head(mtcars)
```

### Quiz

*You can include any number of single or multiple choice questions as a quiz. Use the `question` function to define a question and the `quiz` function for grouping multiple questions together.*

Some questions to verify that you understand the purposes of various base and recommended R packages:

```{r quiz}
quiz(
  question("Which package contains functions for installing other R packages?",
    answer("base"),
    answer("tools"),
    answer("utils", correct = TRUE),
    answer("codetools")
  ),
  question("Which of the R packages listed below are used to create plots?",
    answer("lattice", correct = TRUE),
    answer("tools"),
    answer("stats"),
    answer("grid", correct = TRUE)
  )
)
```

